skater: An R package for SNP-based Kinship Analysis, Testing, and Evaluation

### 2021-07-21 paper/paper.Rmd
affiliations

address:
  - code: 1
    address: Signature Science, LLC., Charlottesville, Virginia, 22911, USA.
  - code: 2
    address: Center for Human Identification, University of North Texas Health Science Center, 3500 Camp Bowie Blvd., Fort Worth, TX, 76107, USA
  - code: 3
    address: Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, 3500 Camp Bowie Blvd., Fort Worth, TX, 76107, USA

@@

Once estimated kinship is converted to degree, it may be of interest to compare the inferred degree to known degrees of relatedness. When aggregated over many relationships and inferences, this can help benchmark performance of a particular kinship analysis method. The skater package adapts a `confusion_matrix()` function from @clark2021 to provide standard contingency table metrics (e.g. sensitivity, specificity, PPV, precision, recall, F1, etc.) with a new reciprocal RMSE (R-RMSE) metric. The R-RMSE metric is defined more thoroughly in the skater package vignette (see Supplementary Material) and may be a preferable measure of classification accuracy when benchmarking relationship degree estimation. Taking the reciprocal of the target and predicted degree in a typical RMSE calculation results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second-degree) than misclassifications at more distant relationships (e.g., classifying a fourth-degree relative pair as fifth-degree).

>>>>

address:
  - code: 1
    address: Signature Science, LLC., Austin, TX 78759, USA.
  - code: 2
    address: Center for Human Identification, University of North Texas Health Science Center, Fort Worth, TX 76107, USA
  - code: 3
    address: Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, Fort Worth, TX 76107, USA

@@

Once estimated kinship is converted to degree, it may be of interest to compare the inferred degree to known degrees of relatedness. When aggregated over many relationships and inferences, this can help benchmark performance of a particular kinship analysis method. The skater package adapts a `confusion_matrix()` function from @clark2021 to provide standard contingency table metrics (e.g. sensitivity, specificity, PPV, precision, recall, F1, etc.) with a new reciprocal RMSE (R-RMSE) metric. The R-RMSE metric is defined more thoroughly in the skater package vignette (see Supplementary Material) and may be a preferable measure of classification accuracy when benchmarking relationship degree estimation. In many kinship benchmarking analyses, classification error is treated in a categorical manner (exact match plus or minus one degree), neglecting the true amount of sharing as a real number. Taking the reciprocal of the target and predicted degree in a typical RMSE calculation results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second-degree) than misclassifications at more distant relationships (e.g., classifying a fourth-degree relative pair as fifth-degree).

### 2021-07-21 paper/paper.tex
affiliations

\address{
\textsuperscript{1}Signature Science, LLC., Charlottesville, Virginia,
22911, USA.\\
\textsuperscript{2}Center for Human Identification, University of North
Texas Health Science Center, 3500 Camp Bowie Blvd., Fort Worth, TX,
76107, USA\\
\textsuperscript{3}Department of Microbiology, Immunology, and Genetics,
University of North Texas Health Science Center, 3500 Camp Bowie Blvd.,
Fort Worth, TX, 76107, USA\\
}

>>>>

\address{
\textsuperscript{1}Signature Science, LLC., Austin, TX 78759, USA.\\
\textsuperscript{2}Center for Human Identification, University of North
Texas Health Science Center, Fort Worth, TX 76107, USA\\
\textsuperscript{3}Department of Microbiology, Immunology, and Genetics,
University of North Texas Health Science Center, Fort Worth, TX 76107,
USA\\
}
